# Mortality improvement

Note: 

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("data-raw", "libraries.r"))

draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")

frs_constants <- readRDS(fs::path(dfrs, "frs_constants.rds"))


```


## Get and extend mortality improvement table

Get SOA MP-2018 mortality improvement table and, using `pendata::extend_mp`:

-   Extend it to earlier ages

-   Extend it to later years

-   Calculate cumulative mortality improvement

-   Adjust cumulative improvement to be indexed to a baseyear

```{r}
#| label: get-extend-mp

# get SOA MP-2018 mortality improvement table

mp1 <- pendata::mp2018
skim(mp1) # 1951-2034, ages 20-120

# for FRS, extend years to 2154, age down to 18, and index to baseyear 2010 ---
mp2 <- pendata::extend_mp(mp1, startage=18, endyear=2154, baseyear=2010) |>
  mutate(system="frs") |>
  relocate(system)
skim(mp2)
ht(mp2)

# mp2 |>
#   filter(year %in% 2009:2011, gender=="female", age %in% 18:21)
# 
# glimpse(mp2)
# ht(mp2)
# count(mp2, gender)
# count(mp2, age) |> ht()
# count(mp2, year) |> ht()

saveRDS(mp2, path(dfrs, "mortality_improvement.rds"))

```

## Show extended mortality improvement table

Here is the result. Show just one gender and age, to fit within what's allowed for client-side data.

```{r}
#| label: show-salary-growth

mp <- readRDS(fs::path(dfrs, "mortality_improvement.rds"))

mp |> 
  filter(gender=="female", age==30) |> 
  DT::datatable(
    options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
    filter="top") |>
  formatPercentage(columns = c("mp"), digits=2) |> 
  formatRound(columns = c("mpc", "mpcadj"), digits=4)

# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)


```


## Comparison to the Reason approach


