# Retiree distribution

Note: 

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("data-raw", "libraries.r"))

draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")

frs_constants <- readRDS(fs::path(dfrs, "frs_constants.rds"))

FileName <- "Florida FRS inputs.xlsx"
fullpath <- fs::path(dfrs, FileName)

```


## Get and save retiree distribution data

```{r}
#| label: get-retiree_distribution

retiree_distribution <- read_excel(fullpath, sheet = "Retiree Distribution")

# glimpse(retiree_distribution)
# 
# skim(retiree_distribution)
# retiree_distribution |>
#   summarise(across(-age, sum))

saveRDS(retiree_distribution, path(dfrs, "retiree_distribution.rds"))


```


## Show retiree_distribution table

Here is the result.

```{r}
#| label: show-salary-growth

rd <- readRDS(fs::path(dfrs, "retiree_distribution.rds"))

rd |> 
  DT::datatable(
          options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
          filter="top") |>
  formatRound(columns = c("n_retire", "total_ben", "avg_ben"), digits=0) |> 
  formatRound(columns = c("n_retire_ratio", "total_ben_ratio"), digits=4)

# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```

## Comparison to the Reason approach

`pendata` constructs the same salary-growth numbers as Reason, but creates a long file rather than a wide file.

Reason reads salary growth rates as a wide file, with a column for each class (Florida FRS model input.R line 180)

Extends salary growth to the maximum yos by carrying the last yos growth rate (yos=30) forward to all subsequent yos (up to 70), and also calculates cumulative growth from yos=0 to yos=70 (Florida FRS benefit model.R lines 6-9)

Note: Later, when calculating entry salary in the salary_headcount table, Reason creates a temporary version of the salary growth table with data for the class being worked on, but does not change the data.



# About -------------------------------------------------------------------

# Get retiree distribution for FRS, from an Excel workbook (Florida FRS
# inputs.xlsx) that Reason created.


# TODO --------------------------------------------------------------------

# Look for additional cleanup code that Reason has in other files and
# consolidate it here.

# Not sure where these data come from. Find out and document.

# Read as text, convert to desired types.

# Reverse engineer spreadsheet, get source data and calculate annual from that,
# giving us the ability to use different interpolation methods.


# setup -------------------------------------------------------------------

source(here::here("data-raw", "libraries.r"))
draw <- here::here("data-raw")

dfrs <- fs::path(draw, "systems", "frs")

FileName <- "Florida FRS inputs.xlsx"
fullpath  <- fs::path(dfrs, FileName)


# get and save data ----------------------------------------------------------------

retiree_distribution <- read_excel(fullpath, sheet = "Retiree Distribution")

glimpse(retiree_distribution)

skim(retiree_distribution)
retiree_distribution |>
  summarise(across(-age, sum))

saveRDS(retiree_distribution, path(dfrs, "retiree_distribution.rds"))
