# Retirement rates


## TODO


## Setup

```{r}
#| label: setup
#| output: false

source(here::here("data-raw", "libraries.r"))
draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")

dxi <- fs::path(dfrs, "Reports", "extracted inputs")

source(fs::path(dfrs, "functions.R"))

frs_constants <- readRDS(fs::path(dfrs, "frs_constants.rds"))

FileName <- "Florida FRS inputs.xlsx"
fullpath <- fs::path(dfrs, FileName)

```

## Function to combine prefix and suffix vectors in order

Combine paste twos string vectors as follows:

-   if prefixes = c("a", "b", "c") and suffixes = c("1", "2")

-   if we call combine(prefixes, suffixes)

-   the result would be c("a_1", "a_2", "b_1", "b_2", "c_1", "c_2") in that order

```{r}
#| label: functions

combine <- function(prefixes, suffixes){
  # paste two string vectors as follows:
  # for example if prefixes = c("a", "b", "c") and suffixes = c("1", "2")
  # the result would be c("a_1", "a_2", "b_1", "b_2", "c_1", "c_2") in that order
  outer(prefixes, suffixes, FUN = paste, sep = "_") |>
    t() |>
    c()
}

```


## Create lists of column names

Combine column-name prefixes with suffixes (gender) to have column names we will use when we read the Excel tables.

```{r}

# create lists of column name prefixes (gender will be suffix)
normal_pre <- c("regular_k12_instructional", "regular_nonk12_instructional",
  "special_risk_all", "elected_officers", "senior_management")

early_pre <- normal_pre[-1]

drop_pre <- c("regular_k12_instructional", "regular_nonk12_instructional",
              "special_risk_xleo", "special_risk_leo", "other")

gender <- c("female", "male")

normal <- list(fnbase="normal retirement", cols=combine(normal_pre, gender))
early <- list(fnbase="early retirement", cols=combine(early_pre, gender))
drop <- list(fnbase="drop entry", cols=combine(drop_pre, gender))

# inspect the list
rrlist <- list(normal, early, drop)
rrlist[[1]]


```


