# Salary data

Get salary data from `Florida FRS inputs.xlsx`, various sheets, convert to a single long tibble, and save.

## TODO

Consider:

-   breaking entrant_profile out into separate R file

-   separate files for salary_headcount_raw, which just relies on one sheet, and salary_headcount, which requires salary_headcount_raw plus salary_growth

-   deleting the NA records

-   interpolation / spline to get every year


## Setup

```{r}
#| label: setup
#| output: false

source(here::here("data-raw", "libraries.r"))
draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")

source(fs::path(dfrs, "functions.R"))



frs_constants <- readRDS(fs::path(dfrs, "frs_constants.rds"))

FileName <- "Florida FRS inputs.xlsx"
fullpath <- fs::path(dfrs, FileName)

```

## Get vector of salary sheetnames

```{r}
#| label: salary-sheetnames

tabs <- excel_sheets(fullpath)
tabs |> str_subset("Regular")
tabs |> str_subset("Salary Distribution")

saltabs <- tabs |>
  str_subset("Salary Distribution ")

```

## Function to get a salary table
```{r}
#| label: salary-function

get_salaries <- function(salsheet){
  # read a salary sheet, pivot and clean, return a tibble

  class_sheetname <- str_remove(salsheet, "Salary Distribution ") # notice ending space
  class <- get_class_from_sheetname(class_sheetname)

  print(paste0("Getting sheet for class: ", class))

  sal1 <- read_excel(fullpath, salsheet, col_types = "text")
  sal2 <- sal1 |>
    pivot_longer(-age, names_to = "yos", values_to = "salary") |>
    mutate(
      class = class,
      age = as.integer(age),
      yos = as.integer(yos),
      salary = as.numeric(salary)
    ) |>
    relocate(class)
  sal2
}


```



## Get all of the salary tables, create a long tibble, and save as rds file


```{r}
#| label: get-and-save-salaries

salary_table <- purrr::map(saltabs, get_salaries) |>
  list_rbind() |>
  mutate(system="frs") |>
  relocate(system) |>
  arrange(system, class, age, yos)

saveRDS(salary_table, fs::path(dfrs, "salary.rds"))

```



## Show salary data

Here is the result.

```{r}
#| label: show-salary-data

sal <- readRDS(fs::path(dfrs, "salary.rds"))

# names(ifd)

DT::datatable(sal,
          options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
          filter="top") |> 
  formatCurrency(columns = c("salary"), currency="", digits=0)


# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)


```


## Comparison to the Reason approach

<!-- 
# Truong's code below, commented out ----------------------------

# Truong gets:
#   salary growth
#   headcount tables
#   salary tables
#   entrant profiles

# He processes eco, eso, and judges separately because "acfr does not provide detailed headcounts"
#   note that the headcounts come from appendix c of the AV, not the ACFR
#   could the so-called less detail simply reflect (1) older start ages for judges, plus
#  (2) either mandatory retirement, or newer plans meaning they don't yet have longer yos???

# Truong deals with this by calculating a special eco eso judges adjustment ratio as the
#   ratio of total eej active members to the sum of active members in the headcount tables
#   for grossing up cell counts so that they hit totals that is different from the gross up he uses
#   for regular, special, etc.; I don't yet know why



# salary_growth_table_ <- read_excel(FileName, sheet = "Salary Growth")
#
# regular_salary_table_ <- read_excel(FileName, sheet="Salary Distribution Regular")
# regular_headcount_table_ <- read_excel(FileName, sheet="HeadCount Distribution Regular") %>%
#   mutate(across(everything(), ~replace(.x, is.na(.x), 0)))
#
# regular_salary_headcount_table <- get_salary_headcount_table(
#   regular_salary_table_,
#   regular_headcount_table_,
#   salary_growth_table,
#   "regular")$salary_headcount_table
#
# regular_entrant_profile_table <- get_salary_headcount_table(
#   regular_salary_table_,
#   regular_headcount_table_,
#   salary_growth_table, "regular")$entrant_profile

# after getting the raw data, Truong's key function is get_salary_headcount_table
# he calls it as follows:
#   regular - 2x returns salary_headcount_table, entrant_profile
#   same for special, admin, eco, eso, judges, senior_management


# Joining headcount data, salary data, and salary growth data

# We account for the Investment Plan (DC plan) head count by inflating the DB head count by the ratio of total system head count to DB head count

# ECO, ESO, and Judges head counts are processed separately as the ACFR does not provide detailed head counts for these classes
# eco_eso_judges_active_member_adjustment_ratio <- eco_eso_judges_total_active_member_ / sum(eco_headcount_table_[-1] + eso_headcount_table_[-1] + judges_headcount_table_[-1])

# total active members (for grossing up) are defined in lines 160+ of Florida FRS model input.R
# they come from numbered page 163 of the ACFR
#Below are the total membership numbers for each membership class. These numbers are from the ACFR and include both DB and DC membership.
# regular_total_active_member_ <- 537128
# special_total_active_member_ <- 72925
# admin_total_active_member_ <- 104
# eco_eso_judges_total_active_member_ <- 2075
# senior_management_total_active_member_ <- 7610



# get_salary_headcount_table <- function(salary_table, headcount_table, salary_growth_table, class_name){
#
#   class_name <- str_replace(class_name, " ", "_")
#
#   if (!class_name %in% c("eco", "eso", "judges")) {
#     assign("total_active_member", get(paste0(class_name, "_total_active_member_")))
#   } else {
#     assign("total_active_member", get("eco_eso_judges_total_active_member_"))
#   }
#
#   salary_growth_table <- salary_growth_table %>%
#     select(yos, contains(class_name)) %>%
#     rename(cumprod_salary_increase = 2)
#
#   salary_table_long <- salary_table %>%
#     pivot_longer(cols = -1, names_to = "yos", values_to = "salary")
#
#   headcount_table_long <- headcount_table %>%
#     pivot_longer(cols = -1, names_to = "yos", values_to = "count") %>%
#     mutate(
#       active_member_adjustment_ratio = if_else(str_detect(class_name, "eco|eso|judges"), eco_eso_judges_active_member_adjustment_ratio,
#                                                total_active_member / sum(count, na.rm = T)),
#       count = count * active_member_adjustment_ratio
#     ) %>%
#     select(-active_member_adjustment_ratio)
#
#   salary_headcount_table <- salary_table_long %>%
#     left_join(headcount_table_long) %>%
#     mutate(
#       yos = as.numeric(yos),
#       start_year = start_year_,
#       entry_age = age - yos,
#       entry_year = start_year - yos) %>%
#     filter(!is.na(salary), entry_age >= 18) %>%
#     left_join(salary_growth_table) %>%
#     mutate(entry_salary = salary / cumprod_salary_increase) %>%
#     select(entry_year, entry_age, age, yos, count, entry_salary)
#
#   entrant_profile <- salary_headcount_table %>%
#     filter(entry_year == max(entry_year)) %>%
#     mutate(entrant_dist = count/sum(count)) %>%
#     select(entry_age, entry_salary, entrant_dist) %>%
#     rename(start_sal = entry_salary)
#
#   output <- list(
#     salary_headcount_table = salary_headcount_table,
#     entrant_profile = entrant_profile)
#
#   return(output)
# }

-->

