# Salary growth rates

Note: Extending salary growth table requires access to frs_constants.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("data-raw", "libraries.r"))

draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")

frs_constants <- readRDS(fs::path(dfrs, "frs_constants.rds"))

FileName <- "Florida FRS inputs.xlsx"
fullpath <- fs::path(dfrs, FileName)

```


## Get salary growth table

```{r}
#| label: get-salary-growth

salgrowth1 <- read_excel(fullpath, sheet = "Salary Growth", col_types = "text")

salgrowth2 <- salgrowth1 |>
  pivot_longer(-yos, names_to = "class", values_to = "salgrowth") |>
  mutate(
    yos = as.integer(yos),
    class = str_remove(class, "salary_increase_"),
    # fix inconsistent naming
    class = ifelse(class == "special_risk", "special", class),
    salgrowth = as.numeric(salgrowth),
    system="frs"
  ) |>
  select(system, class, yos, salgrowth)

glimpse(salgrowth2)
# skim(salgrowth2)
# ht(salgrowth2)


```


## Extend salary growth table to max yos and calculate cumulative growth

Extends salary growth to the maximum yos by carrying the last yos growth rate (yos=30) forward to all subsequent yos (up to 70), and calculates cumulative growth from yos=0 to yos=70 (same as in Florida FRS benefit model.R lines 6-9).


```{r}
#| label: extend-salary-growth

# extend salary growth table to max yos and calculate cumulative growth -------
# cumprod(1 + lag(.x, default = 0))
# salgrowth2 <- readRDS(path(dfrs, "salary_growth_raw.rds"))

salgrowthx <- crossing(salgrowth2 |>
                         select(system, class),
                       yos=0:frs_constants$yos_max) |>
  left_join(salgrowth2, by = join_by(system, class, yos)) |>
  arrange(system, class, yos) |>
  group_by(system, class) |>
  fill(salgrowth, .direction = "down") |>
  mutate(cumprod_increase=cumprod(1 + lag(salgrowth, default = 0))) |>
  ungroup()

saveRDS(salgrowthx, fs::path(dfrs, "salary_growth.rds"))
```


## Show final salary growth table

Here is the result.

```{r}
#| label: show-salary-growth

salgrowthx <- readRDS(fs::path(dfrs, "salary_growth.rds"))

DT::datatable(salgrowthx,
          options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
          filter="top") |>
  formatPercentage(columns = c("salgrowth"), digits=2) |> 
  formatRound(columns = c("cumprod_increase"), digits=4)

# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```

## Comparison to the Reason approach

`pendata` constructs the same salary-growth numbers as Reason, but creates a long file rather than a wide file.

Reason reads salary growth rates as a wide file, with a column for each class (Florida FRS model input.R line 180)

Extends salary growth to the maximum yos by carrying the last yos growth rate (yos=30) forward to all subsequent yos (up to 70), and also calculates cumulative growth from yos=0 to yos=70 (Florida FRS benefit model.R lines 6-9)

Note: Later, when calculating entry salary in the salary_headcount table, Reason creates a temporary version of the salary growth table with data for the class being worked on, but does not change the data.

