---
format: html
editor-options: 
  chunk-output-type: console
---

# Texas TRS

## Tables needed

-   salary_headcount_table
-   term_rate_age
-   mort_table
-   entrant_profile_table
-   mort_retire_table
-   normal_retire_rate_table
-   early_retire_rate_table
-   separation_rate_table
-   salary_growth_table

## Setup

```{r}
#| label: libraries

source(here::here("data-raw", "txtrs", "libraries.r"))
# search() # list attached packages and R objects
# sessioninfo::session_info()
# library(pendata)

```

```{r}
#| label: constants-and-folders

system_name <- "txtrs"

ddata <- here::here("data")
draw <- here::here("data-raw")
dtxtrs <- fs::path(draw, "txtrs")
dinputs <- fs::path(dtxtrs, "Inputs")

fname <- "TxTRS_BM_Inputs.xlsx"
fpath <- fs::path(dtxtrs, fname)

```

## Overview

Create a set of lists that will be elements in a master list called txtrs

```{r}
#| label: start-txtrs

txtrs <- list()
```


## 1. Actuarial and economic assumptions

```{r}
#| label: actuarial

rates <- list()
rates <- within(rates, {
  dr_current_ <- 0.07 # discount rate for current members
  dr_new_ <- 0.07 # discount rate for new members
  payroll_growth_ <- 0.029 # payroll growth assumption
  pop_growth_ <- 0 # plan's active population growth assumption
  inf_rate <- 0.023 # inflation rate assumption
  workforce_scenario_ <- "baseline"
})


```

```{r}
#| label: mortality

mort <- list()
mort$ScaleMultiple <- 0.9

```

# 2. Benefit assumptions

```{r}
#| label: defined-benefit

#.. Defined benefit assumptions: ----
db <- list()
db <- within(db, {
  ReduceRetAge <- 55
  EarlyRetYOS <- 30
  NormalRetAgeI <- 65
  NormalYOSI <- 5
  NormalRetRule <- 80
  NormalRetRuleAgeI <- 60
  NormalRetRuleAgeII <- 62

  FinAvgSalaryYears_current <- 5 #final average salary period
  FinAvgSalaryYears_gft <- 3 #final average salary period for grandfathered members

  Interest <- 0.02 #interest on DB employee balance
  BenMult <- 0.023 #DB benefit multiplier
})

```

```{r}
#| label: cola

#.. Cost of living adjustment (COLA) assumptions: ----
cola <- list()
cola <- within(cola, {
  COLA_current_active <- 0 #COLA for current active members
  COLA_new_active <- 0 #COLA for new active members (note that this applies to the cash balance plan too if new members go to the CB plan)
  COLA_current_retire <- 0 #COLA for current retirees
  COLA_current_retire_one <- 0 #One-time COLA for current retirees
  one_time_cola_ <- F #One-time COLA or not? True means yes, False means no
})

```

```{r}
#| label: cash-balance

#.. Cash balance benefit assumptions ----
cashbal <- list()
cashbal <- within(cashbal, {
  CB_EE_paycredit <- 0.06 #Cash balance employee contribution rate
  CB_ER_paycredit <- 0.09 #Cash balance employer pay credit
  CB_vesting <- 5

  icr_smooth_period_ <- 5 #Smoothing period for the interest crediting rate (ICR)
  icr_floor_ <- 0.04 #Minimum ICR granted
  icr_cap_ <- 0.07 #Maximum ICR granted
  icr_upside_share_ <- 0.5 #Upside sharing ratio of the ICR
  icr_return_vol_ <- 0.12 #Assumed annual volatility of the investment returns to be used for the ICR calculation

  ACR <- 0.04 #Annuity Conversion Rate: interest rate to convert the cash balance into retirement annuity
})

```

```{r}
#| label: defined-contribution

#.. DC benefit assumptions ----
dc <- list()
dc <- within(dc, {
  dc_ee_cont_ <- 0.0825 - cashbal$CB_EE_paycredit #Defined contribution employee contribution rate
  dc_arith_return_ <- 0.07 #Arithmetic return assumption for the DC plan
  dc_return_sd_ <- 0.12 #Return volatility assumption for the DC plan
})

```

# 3. Funding assumptions

```{r}
#| label: funding

funding <- list()
funding <- within(funding, {
  DB_EE_cont <- 0.0825
  extra_er_stat_cont_ <- 0 #Additional employer contribution rate for the statutory funding policy
  public_edu_payroll_percent_ <- 0.588 #Percent of total payroll eligible for the Public Education Employer Contribution
  funding_policy_ <- "statutory" #"ADC"

  #.. Amortization policy ----
  amo_pay_growth_ <- 0.029
  amo_period_current_ <- 30
  amo_period_new_ <- 30
  amo_method_ <- "level %"
})

``` 



# 4. Investment assumptions

```{r}
#| label: investment-assumptions

invest <- list()
invest <- within(invest, {
  return_scen_ <- "assumption" #"recession", "recur_recession", "constant_6"
  model_return_ <- 0.07
})

```

# 5. Plan design assumptions
```{r}
#| label: plan-design

design <- list()
design$db_new_ratio_ <- 1 #% of new hires that will join the DB plan.

```

# 6. Model assumptions

```{r}
#| label: model-assumptions

model <- list()
model <- within(model, {
  ModelPeriod <- 30 #Projection period (typically 30 years)
  MinAge <- 20 #Age of the typical youngest member
  MaxAge <- 120 #Max age from mortality assumptions
  YearStart <- 2024 #Year of the latest val report
  MinYear <- 1980 #No hard rule about this. Should get back to about 40 years from now.
  MaxYear <- YearStart + ModelPeriod + MaxAge - MinAge

  EntryYear <- MinYear:(YearStart + ModelPeriod)
  Years <- MinYear:MaxYear
  Age <- MinAge:MaxAge
  YOS <- 0:70
  RetirementAge <- Age

  retiree_pop_current <- 508701
  ben_payment_current <- 15258219146
  retire_refund_ratio_ <- 0.3

  cal_factor_ <- 12.1 / 12.189475 #This is to adjust the DB benefits to result in a PVFB that matches (or mostly matches) the plan's number. Update this when a new valuation report is released.
  nc_cal_ <- 12.1 / 12.1117 #This is to further calibrate the normal cost in the funding model. Update this when a new valuation report is released.
  ben_cal_ <- 1 # 1.06               #This is to calibrate the benefit payments and the PV of future benefits for current retirees. Update this when a new valuation report is released.

  # $10,237,518,335 is the PVFB for current term vested members
  # $6,471,070,551  is the remaining gap between the val's AAL and the model's AAL.
  # Update this when a new valuation report is released.
  PVFB_term_current <- 273095060051 - 262639131631 #Val's AAL - model's AAL
  amo_period_term <- 50
})

```


# 7. Import key data tables

```{r}
#| label: get-data

FileName <- 'TxTRS_BM_Inputs.xlsx'
FileName <- fs::path(dtxtrs, FileName)

#.. survival and mortality rates ----
SurvivalRates <- read_excel(FileName, sheet = 'Mortality Rates')

MaleMP <- read_excel(FileName, sheet = 'MP-2018_Male')
FemaleMP <- read_excel(FileName, sheet = 'MP-2018_Female')

#.. soa survival rates ....
SurvivalRates_ <- read_excel(
  fs::path(dtxtrs, "Inputs/pub-2010-amount-mort-rates.xlsx"),
  sheet = "PubT-2010(B)"
)

male_mp_table_ <- read_excel(
  fs::path(dtxtrs, 'Inputs/mp-2021-rates.xlsx'),
  sheet = "Male"
)

female_mp_table_ <- read_excel(
  fs::path(dtxtrs, 'Inputs/mp-2021-rates.xlsx'),
  sheet = "Female"
)

SalaryGrowthYOS <- read_excel(FileName, sheet = "Salary Growth YOS")
SalaryMatrix <- read_excel(FileName, sheet = "Salary Matrix")
HeadCountMatrix <- read_excel(FileName, sheet = "Head Count Matrix")
SalaryEntry <- read_excel(FileName, sheet = "Entrant Profile")

TerminationRateAfter10 <- read_excel(
  FileName,
  sheet = 'Termination Rates after 10'
)

TerminationRateBefore10 <- read_excel(
  FileName,
  sheet = 'Termination Rates before 10'
)

RetirementRates <- read_excel(FileName, sheet = 'Retirement Rates')

ReducedGFT <- read_excel(FileName, sheet = "Reduced GFT")
ReducedOthers <- read_excel(FileName, sheet = "Reduced Others")

RetireeDistribution <- read_excel(FileName, sheet = "Retiree Distribution")

funding_data <- read_excel(FileName, sheet = "Funding Data")
return_scenarios <- read_excel(FileName, sheet = "Return Scenarios")

```


## Save the txtrs list

```{r}
#| label: save-txtrs

# usethis::use_data(txtrs, overwrite = TRUE)

```

```{r}
#| label: stop-here
knitr::knit_exit()

```


## base mortality table -  use the pub-2010 amount table from SOA ----

```{r}
#| label: base-mortality
#| eval: false

load(here::here("data", "pub2010amount_mortality_rates.rda"))

base_mortality_rates <- pub2010amount_mortality_rates |>
  mutate(system = system_name) |>
  select(system, employee_type, beneficiary_type, gender, age, rate)
glimpse(base_mortality_rates) # stacked version
txtrs$base_mortality_rates <- base_mortality_rates

```


## mortality improvement table

```{r}
#| label: mortality-improvement

# these appear to be soi mp tables?
fpath <- fs::path(dinputs, "mp-2021-rates.xlsx")
male_mp_table_ <- read_excel(fpath, sheet = "Male")
female_mp_table_ <- read_excel(fpath, sheet = "Female")

```



# age beneficiary_type gender rate

# mortality improvement table ----
male_mp_table_ <- read_excel('Inputs/mp-2021-rates.xlsx', sheet = "Male")
female_mp_table_ <- read_excel('Inputs/mp-2021-rates.xlsx', sheet = "Female")


# tx trs tables Reason used ----
fname <- "TxTRS_BM_Inputs.xlsx"
fpath <- fs::path(dtxtrs, fname)

survival_rates <- read_excel(fpath, sheet = "Mortality Rates")

MaleMP <- read_excel(fpath, sheet = "MP-2018_Male")
FemaleMP <- read_excel(fpath, sheet = "MP-2018_Female")

SalaryGrowthYOS <- read_excel(fpath, sheet = "Salary Growth YOS")
SalaryMatrix <- read_excel(fpath, sheet = "Salary Matrix")
HeadCountMatrix <- read_excel(fpath, sheet = "Head Count Matrix")
SalaryEntry <- read_excel(fpath, sheet = "Entrant Profile")

TerminationRateAfter10 <- read_excel(fpath, sheet = 'Termination Rates after 10')
TerminationRateBefore10 <- read_excel(fpath, sheet = 'Termination Rates before 10')
RetirementRates <- read_excel(fpath, sheet = 'Retirement Rates')

ReducedGFT <- read_excel(fpath, sheet = "Reduced GFT")
ReducedOthers <- read_excel(fpath, sheet = "Reduced Others")

RetireeDistribution <- read_excel(fpath, sheet = "Retiree Distribution")

funding_data <- read_excel(fpath, sheet = "Funding Data")
return_scenarios <- read_excel(fpath, sheet = "Return Scenarios")


# DON'T RUN -- Reason code from TxTRS_model_inputs.R ----
#7. Import key data tables
FileName <- 'TxTRS_BM_Inputs.xlsx'

SurvivalRates <- read_excel(FileName, sheet = 'Mortality Rates')
MaleMP <- read_excel(FileName, sheet = 'MP-2018_Male')
FemaleMP <- read_excel(FileName, sheet = 'MP-2018_Female')

SurvivalRates_ <- read_excel("Inputs/pub-2010-amount-mort-rates.xlsx", sheet = "PubT-2010(B)")
male_mp_table_ <- read_excel('Inputs/mp-2021-rates.xlsx', sheet = "Male")
female_mp_table_ <- read_excel('Inputs/mp-2021-rates.xlsx', sheet = "Female")

SalaryGrowthYOS <- read_excel(FileName, sheet = "Salary Growth YOS")
SalaryMatrix <- read_excel(FileName, sheet = "Salary Matrix")
HeadCountMatrix <- read_excel(FileName, sheet = "Head Count Matrix")
SalaryEntry <- read_excel(FileName, sheet = "Entrant Profile")

TerminationRateAfter10 <- read_excel(FileName, sheet = 'Termination Rates after 10')
TerminationRateBefore10 <- read_excel(FileName, sheet = 'Termination Rates before 10')
RetirementRates <- read_excel(FileName, sheet = 'Retirement Rates')

ReducedGFT <- read_excel(FileName, sheet = "Reduced GFT")
ReducedOthers <- read_excel(FileName, sheet = "Reduced Others")

RetireeDistribution <- read_excel(FileName, sheet = "Retiree Distribution")

funding_data <- read_excel(FileName, sheet = "Funding Data")
return_scenarios <- read_excel(FileName, sheet = "Return Scenarios")
