# Headcount tables

## Overview

### Details of Reason approach

Reason reads raw salary-headcount data in `Florida FRS model input.R` lines 182-208. These counts do not include DC plan headcount. Reason adjusts headcount data upward to be consistent with plan totals,

## TODO

-   Near term: Determine precise source of the data. Presumably it is the FRS 2022 actuarial valuation, but we should know the exact table number.


## Setup

```{r}
#| label: setup
#| output: false

draw <- here::here("data-raw")
dfrs <- fs::path(draw, "systems", "frs")
drds <- fs::path(dfrs, "rds")

source(fs::path(draw, "libraries.r"))
source(fs::path(dfrs, "functions.r"))

frs_constants <- readRDS(fs::path(drds, "frs_constants.rds"))

FileName <- "Florida FRS inputs.xlsx"
fullpath <- fs::path(dfrs, FileName)

```


## Function to get a single headcount sheet

```{r}
#| label: headcount-functions

get_headcount <- function(fullpath, hcsheet){
  # read a headcount sheet, pivot and clean, return a tibble

  class_sheetname <- str_remove(hcsheet, "HeadCount Distribution ") # notice ending space
  class <- get_class_from_sheetname(class_sheetname)

  print(class)

  hc1 <- read_excel(fullpath, hcsheet, col_types = "text")
  hc2 <- hc1 |>
    pivot_longer(-age, names_to = "yos", values_to = "count") |>
    mutate(
      class = class,
      age = as.integer(age),
      yos = as.integer(yos),
      count = as.numeric(count)
    ) |>
    relocate(class)
  hc2
}

```


## Review and define desired sheets

```{r}

tabs <- excel_sheets(fullpath)

tabs |> str_subset("Regular")
tabs |> str_subset("HeadCount Distribution") # e.g., HeadCount Distribution Sen Man


# get all of the headcount tables ----
hctabs <- tabs |>
  str_subset("HeadCount")


```



## Get data

```{r}
#| label: get-data

headcount_table <- hctabs |> 
  purrr::map(\(x) get_headcount(fullpath, x)) |>
  list_rbind() |>
  mutate(system="frs") |>
  relocate(system) |>
  arrange(system, class, age, yos)

```

## Gross up the headcount

Headcount does not include DC members, so gross up for that using ratio of total headcount to the DB headcount. Requires `frs_constants$total_actives`.

```{r}
#| label: headcount-grossup


# active_member_adjustment_ratio = if_else(str_detect(class_name, "eco|eso|judges"), eco_eso_judges_active_member_adjustment_ratio,
#                                          total_active_member / sum(count, na.rm = T))

# headcount_table <- readRDS(path(dfrs, "headcount_raw.rds"))

# compute adjustment ratios based on total headcount including DC
adjustment_ratios1 <- headcount_table |>
  mutate(class=ifelse(class %in% c("eco", "eso", "judges"),
                      "eco_eso_judges",
                      class)) |>
  summarise(count=sum(count, na.rm=TRUE), .by=class) |>
  left_join(frs_constants$total_actives, by = join_by(class)) |>
  mutate(active_member_adjustment_ratio =
           total_actives / count)

# break eco_eso_judges into 3 classes
adjustment_ratios2 <- adjustment_ratios1 |>
  select(class, active_member_adjustment_ratio) |>
  mutate(reps = ifelse(class=="eco_eso_judges", 3, 1)) |>
  uncount(reps) |>
  mutate(class = ifelse(class=="eco_eso_judges",
                        c("eco", "eso", "judges"),
                        class))
# adjustment_ratios2

headcount_adjusted <- headcount_table |>
  left_join(adjustment_ratios2, by = join_by(class)) |>
  mutate(count=count * active_member_adjustment_ratio) |>
  select(-active_member_adjustment_ratio)

# headcount_adjusted

saveRDS(headcount_adjusted, path(drds, "headcount.rds"))

```



## Show the final headcount

```{r}
#| label: show-final-headcount-data

hca <- readRDS(fs::path(drds, "headcount.rds"))

hca |> 
  DT::datatable(
    options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
    filter="top") |> 
  formatRound(columns = c("count"), digits=0)


# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```
