# Retiree distribution

Read retiree distribution data from "Florida FRS inputs.xlsx", sheet = "Retiree Distribution". 

We do not yet know precisely where these data come from. It looks like Reason obtained data on the number of retirees and total benefits by 5-year age groups (presumably from the 2022 actuarial valuation) and did a simple division by 5 to spread them out equally among individual years of age.

## TODO

-   Near term: Determine precise source of the data. Presumably it is the FRS 2022 actuarial valuation, but we should know the exact table number.

-   Longer term: Consider a spline approach for distributing 5-year age-group data to individual ages. See [pendata issue #4](https://github.com/donboyd5/pendata/issues/4).

## Setup

```{r}
#| label: setup
#| output: false

source("../_common_frs.R")

```


## Get and save retiree distribution data

The columns n_retire_ratio and total_ben_ratio give each single-year of age's share of total retirees and total benefits, respectively. Thus, these columns sum to 1.


```{r}
#| label: get-retiree_distribution

retiree_distribution <- read_excel(fullpath, sheet = "Retiree Distribution")

# glimpse(retiree_distribution)
# 
# skim(retiree_distribution)

# verify that n_retire_ratio and total_ben_ratio sum to one
# retiree_distribution |>
#   select(-avg_ben) |> 
#   summarise(across(-age, sum))

saveRDS(retiree_distribution, path(drds, "retiree_distribution.rds"))


```


## Show retiree_distribution table

Here is the result.

```{r}
#| label: show-retiree-distribution

rd <- readRDS(fs::path(drds, "retiree_distribution.rds"))

rd |> 
  DT::datatable(
          options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
          filter="top") |>
  formatRound(columns = c("n_retire", "total_ben", "avg_ben"), digits=0) |> 
  formatRound(columns = c("n_retire_ratio", "total_ben_ratio"), digits=4)

# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```

## Comparison to the Reason approach

Reason does the same thing in `Florida FRS model input.R`, line 245,

