---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Retirement rates


## TODO


## Setup

```{r}
#| label: setup
#| output: false

# source("/data-raw/frs/_common_frs.R")
source(here::here("data-raw", "frs", "_common_frs.R"))

```

## Function to combine prefix and suffix vectors in order

Function `combine` combines two string vectors as follows:

-   if prefixes = c("a", "b", "c") and suffixes = c("1", "2")

-   if we call combine(prefixes, suffixes)

-   the result is c("a_1", "a_2", "b_1", "b_2", "c_1", "c_2") in that order

```{r}
#| label: functions

combine <- function(prefixes, suffixes, sep="_"){
  # paste two string vectors as follows:
  # for example if prefixes = c("a", "b", "c") and suffixes = c("1", "2")
  # the result would be c("a_1", "a_2", "b_1", "b_2", "c_1", "c_2") in that order
  outer(prefixes, suffixes, FUN = paste, sep = sep) |>
    t() |>
    c()
}

```


## Create lists of column names

Combine column-name prefixes with suffixes (gender) to have column names we will use when we read the Excel tables.

```{r}
#| label: create-column-name-lists

# create lists of column name prefixes (gender will be suffix)
normal_pre <- c("regular_k12_instructional", "regular_nonk12_instructional",
  "special_risk_all", "elected_officers", "seniormanagement") # note senior.management is 1 word

early_pre <- normal_pre[-1]

drop_pre <- c("regular_k12_instructional", "regular_nonk12_instructional",
              "special_risk_nonleo", "special_risk_leo", "other")

gender <- c("female", "male")

normal <- list(fnbase="normal retirement", cols=combine(normal_pre, gender))
early <- list(fnbase="early retirement", cols=combine(early_pre, gender))
drop <- list(fnbase="drop entry", cols=combine(drop_pre, gender))

# inspect the list
rrlist <- list(normal, early, drop)
rrlist[[1]]
# rrlist[[3]]

```

## Function to get tier1 and tier2 retirement rate tables from Excel

tier 1 and tier 2 are in separate sheets. The function reads both sheets and stacks the results

```{r}
#| label: functions-tiers

get_tier <- function(fpath, colnames){
  # get a single tier from an excel sheet and clean it up
  print(fpath)
  
  # base file name without extension
  fnbase <- fpath |> 
    fs::path_file() |> 
    fs::path_ext_remove()

  # get everything before "tier"
  retirement_type <- str_extract(fnbase, ".*(?=tier)") |> 
    str_trim() |> 
    str_replace(" ", "_")
  
  # get everything from "tier" to the end
  tier <- str_extract(fnbase, "tier.*") |> 
    str_remove(" ")

  tier_df <- read_excel(fpath,
                    sheet = 1,
                    range = cellranger::cell_cols(1:length(colnames)),
                    col_names = colnames,
                    col_types = "text")

  # we only want rows for which age is good
  first_data_row <- which(tier_df$agec=="Age")[1] + 1

  # last data row is last not-na row after first data row
  na_indices <- which(is.na(tier_df$agec))
  last_data_row <- na_indices[na_indices > first_data_row][1] - 1

  tier_df <- tier_df |>
    slice(first_data_row:last_data_row) |>
    mutate(retirement_type=retirement_type, tier=tier) |>
    pivot_longer(cols=-c(retirement_type, tier, agec),
                 values_to = "rate") |>
    filter(!is.na(rate)) |>
    select(retirement_type, tier, name, agec, rate)

  tier_df
}

get_retrates <- function(rrpath, fnbase, cols){
  
  # given a base file name and set of columns that are common to two tiers,
  # e.g., "normal retirement"
  # call get_tier for each tier and combine the resulting data frames
  
  fnames <- paste0(fnbase, c(" tier 1", " tier 2"), ".xlsx")
  fpaths <- fs::path(rrpath, fnames)
  tier1_path <- fpaths[1]
  tier2_path <- fpaths[2]
  colnames <- c("agec", cols) # agec is age as character

  bind_rows(get_tier(tier1_path, colnames), 
            get_tier(tier2_path, colnames))
}

```

## Get and save raw retirement rates

```{r}
#| label: get-save-raw-rates

# tmp <- get_retrates(rrlist[[2]])
# count(tmp, agec)

rrates_raw <- rrlist |>
  purrr::map(\(x) get_retrates(rrpath=dxi, fnbase=x$fnbase, cols=x$cols)) |>
  list_rbind() |>
  separate(name, into = c("vname", "gender"), sep = "_(?=[^_]+$)", remove = FALSE)

# glimpse(rrates_raw)
# count(rrates_raw, agec)
count(rrates_raw, retirement_type)

saveRDS(rrates_raw, path(drds, "rrates_raw.rds"))
```


## Replace each summary age 70-79 record with one record for individual age

-   All retirement types and tiers: Create a new row for each individual age from 70-79 that has the retirement rate for the 70-79 age group, in place of the grouped row (Reason "Florida FRS benefit model.R" lines 300-323, clean_retire_rate_table())


```{r}
#| label: add-records-7079

rr1 <- readRDS(path(drds, "rrates_raw.rds"))
ht(rr1)
rr1$agec |> unique() |> sort()

# fill in the 70-79 age group by repeating 70 - interpolation probably better
rr2 <- rr1 |>
  mutate(reps=ifelse(agec=="70-79", 10, 1)) |>
  uncount(reps) |> # 10 copies of each 70-79 record
  mutate(age=as.numeric(agec),
         rate=as.numeric(rate)) |>
  mutate(age=ifelse(agec=="70-79", 70:79, age), .by=c(retirement_type, tier, vname, gender)) |>
  select(-agec) |> 
  arrange(retirement_type, tier, name, age)

```


## Add zero-records for ages 45-49 for normal retirement tier 2

-   Normal retirement tier 2: Create new rows for individual ages 45-49 with zero retirement rate (this matches the min retirement age in the data for normal tier 1) (Reason "Florida FRS benefit model.R" lines 353-355) **What's the general rule here? Do we want retirement rates starting at age 45 for all retirement types and tiers?**


```{r}
#| label: add-records-normal-4549

rr2 |>
  summarise(minage=min(age), maxage=max(age), .by=c(retirement_type, tier))

new_rows <- rr2 |> 
  filter(retirement_type=="normal_retirement", tier=="tier2") |> 
  select(-c(age, rate)) |> 
  distinct() |> # we don't want all ages
  mutate(reps=5) |> 
  uncount(reps) |>
  mutate(age=45:49,
         rate=0, .by=c(retirement_type, tier, name))

# count(new_rows, age)
# count(new_rows, retirement_type, tier, name)
# glimpse(new_rows)

rr3 <- bind_rows(rr2, new_rows) |> 
  arrange(retirement_type, tier, name)

# skim(rr3)

# the result
rr3 |>
  summarise(minage=min(age), maxage=max(age), .by=c(retirement_type, tier))

```

## Replace special risk DROP entry records with average of LEO and non-LEO values

See "Florida FRS benefit model.R" lines 358-372.

```{r}
#| label: special-risk-DROP-combine

to_replace <- expression(retirement_type=="drop_entry" & str_detect(name, "special")) # boolean

replacement_recs <- rr3 |> 
  filter(eval(to_replace)) |> 
  summarise(rate=mean(rate), .by=c(retirement_type, tier, gender, age)) |> 
  mutate(vname="special_risk") |> 
  unite(name, vname, gender, remove = FALSE)

# glimpse(replacement_recs)
# count(replacement_recs, gender)
# count(replacement_recs, age)
# skim(replacement_recs)

rr4 <- rr3 |> 
  filter(!eval(to_replace)) |> 
  bind_rows(replacement_recs) |> 
  arrange(retirement_type, tier, name, vname, gender, age) |> 
  relocate(rate, .after = age)

```


## Compute summary retirement rates for each group and tier

This translates get_normal_retire_rate_table()

See "Florida FRS benefit model.R" lines 374-453.

Reason gets a single retirement rate by age, for each class-tier group.

For group and age:

-   First, for each employee_type Reason adds retirement + drop-entry rates.
-   The single rate is then the mean of these sums.


```{r}
#| label: compute-summary-retirement-rates

rr5 <- rr4 |> 
  mutate(group=str_split(string = name, pattern = "_", n = 2, simplify = TRUE)[, 1]) |> 
  summarise(rate=sum(rate), .by=c(group, tier, vname, gender, age)) |> 
  summarise(rate=mean(rate), .by=c(group, tier, age))

count(rr5, group, tier)
# rr5 is the base we use to create full retirement rate tables for all classes

```

## Final file

Match retirement rates with classes, retirement_types, and tiers to create a file that has a single retirement rate for each class, retirement_type, tier, and age combination.

```{r}
#| label: final-file

# frs_constants$classes

class_group_xwalk <- tribble(
  ~class, ~normal_retirement, ~drop_entry,
  "regular", "regular", "regular",
  "special", "special", "special",
  "admin", "special", "special",
  "eco", "elected", "other",
  "eso", "elected", "other",
  "judges", "elected", "other",
  "seniormanagement", "seniormanagement", "other"
  ) |> 
  pivot_longer(cols = -class, 
               names_to = "retirement_type",
               values_to = "group")

rr6 <- class_group_xwalk |> 
  right_join(rr5,
             by = join_by(group),
             relationship = "many-to-many")

# count(rr6, class)
# count(rr6, retirement_type)
# count(rr6, tier)
# count(rr6, class, retirement_type, tier)

```


## Save final retirement rates file


```{r}
#| label: save-retirement-rates

saveRDS(rr6, path(drds, "retirement_rates.rds"))

```



## Reason code

```{r}
#| label: truong-special-risk-DROP
#| eval: false

special_risk_drop_entry_tier_1_table <- drop_entry_tier_1_table %>% 
  select(age, contains("special_risk")) %>% 
  mutate(
    special_risk_female = (special_risk_non_leo_female + special_risk_leo_female)/2,
    special_risk_male = (special_risk_non_leo_male + special_risk_leo_male)/2,
    .keep = "unused"
    )

special_risk_drop_entry_tier_2_table <- drop_entry_tier_2_table %>% 
  select(age, contains("special_risk")) %>% 
  mutate(
    special_risk_female = (special_risk_non_leo_female + special_risk_leo_female)/2,
    special_risk_male = (special_risk_non_leo_male + special_risk_leo_male)/2,
    .keep = "unused"
  )

```



```{r}
#| label: truong-get_normal_retire_rate_table
#| eval: false

  
get_normal_retire_rate_table <- function(class_name, drop_entry_table, normal_retire_rate_table){
  
  if (class_name %in% c("eco", "eso", "judge")) {
    search_text_in_drop_entry_table = "other"
    search_text_in_normal_retire_table = "eco_eso_jud"
  } else if (class_name == "senior management") {
    search_text_in_drop_entry_table = "other"
    search_text_in_normal_retire_table = "senior_management"
  } else if (class_name == "admin"){
    search_text_in_drop_entry_table = search_text_in_normal_retire_table = "special"
  } else {
    search_text_in_drop_entry_table = search_text_in_normal_retire_table = class_name
  }
  
  normal_retire_rate_table <- (drop_entry_table %>% select(contains(search_text_in_drop_entry_table)) + 
                                 normal_retire_rate_table %>% select(contains(search_text_in_normal_retire_table))) %>% 
    add_column(age = drop_entry_table$age, .before = 1) %>% 
    rowwise() %>% 
    mutate(normal_retire_rate = mean(c_across(-age)), .keep = "unused") %>% 
    ungroup()
  
  return(normal_retire_rate_table)
}

regular_normal_retire_rate_tier_1_table <- get_normal_retire_rate_table(class_name = "regular",
                                                                        drop_entry_table = drop_entry_tier_1_table,
                                                                        normal_retire_rate_table = normal_retire_rate_tier_1_table)

...

```














<!-- Retirement rate tables source: Florida FRS model input.R -->

<!-- drop_entry_tier_1_table_ <- read_excel("Reports/extracted inputs/drop entry tier 1.xlsx") -->
<!-- drop_entry_tier_2_table_ <- read_excel("Reports/extracted inputs/drop entry tier 2.xlsx") -->

<!-- normal_retirement_tier_1_table_ <- read_excel("Reports/extracted inputs/normal retirement tier 1.xlsx") -->
<!-- normal_retirement_tier_2_table_ <- read_excel("Reports/extracted inputs/normal retirement tier 2.xlsx") -->

<!-- early_retirement_tier_1_table_ <- read_excel("Reports/extracted inputs/early retirement tier 1.xlsx") -->
<!-- early_retirement_tier_2_table_ <- read_excel("Reports/extracted inputs/early retirement tier 2.xlsx") -->

<!-- source: Florida FRS benefit model.R -->
<!-- clean_retire_rate_table <- function(df, col_names){ -->

<!--   index_of_na_row_from_bottom <- tail(which(rowSums(is.na(df)) == ncol(df)),1) -->
<!--   index_of_row_before_body <- which(df[,1] == "Age") -->

<!--   df <- df %>% -->
<!--     slice(-(index_of_na_row_from_bottom:n())) %>% -->
<!--     slice(-(1:index_of_row_before_body)) %>% -->
<!--     select_if(~any(!is.na(.))) -->

<!--   index_of_70_79_row <- which(df[,1] == "70-79") -->

<!--   names(df) <- col_names -->

<!--   df <- df %>% -->
<!--     add_row(age=as.character(71:79), .after=index_of_70_79_row) %>% -->
<!--     mutate( -->
<!--       age = replace(age, age == "70-79", "70"), -->
<!--       across(everything(), ~as.numeric(.x)) -->
<!--     ) %>% -->
<!--     fill(everything(), .direction = "down") -->

<!--   return(df) -->
<!-- } -->
