# Salary and headcount tables combined, and extract entrants

Combine salary and headcount tables, and compute entry salary:

-   Get the separate salary and headcount tables (which have 5-year age and yos groups)
-   Get the salary growth table
-   Merge headcount and salary
-   Compute entry age and entry year
-   Merge salary growth
-   Compute implied entry-year salary from cumulative salary growth
-   Save salary-headcount table

Reason does much of this near and before `Florida FRS model input.R` lines
  182-208. (TODO: Improve this documentation.)

Entrant profile:

-   From salary-headcount table get most-recent entry year
-   For each class of employee, compute \# of workers in each entry-age group as share of all entrants in the class
-   Keep class, entry age group (entry_age), start_sal, and \# of entrants in entry_age group as share of all entrants in the class
-   Save entrant profile

Reason constructs the entrant profile in `Florida FRS benefit model.R` lines 55-59.

## TODO

## Setup

```{r}
#| label: setup
#| output: false

source("../_common_frs.R")

```

## Get input data

```{r}
#| label: get-data


sg1 <- readRDS(path(drds, "salary_growth.rds"))
sal1 <- readRDS(path(drds, "salary.rds"))
hc1 <- readRDS(path(drds, "headcount.rds")) # grossed up


```

## Create salary-headcount table

```{r}
#| label: salary-headcount

salary_headcount <- hc1 |>
  left_join(sal1, by = join_by(system, class, age, yos)) |>
  mutate(start_year = frs_constants$start_year,
         entry_age = age - yos,
         entry_year = start_year - yos) |>
  # why do we have unneeded years? maybe weed out earlier
  filter(!is.na(salary), entry_age >= 18) |>  
  left_join(sg1, by = join_by(system, class, yos)) |>
  mutate(entry_salary = salary / cumprod_increase) |>
  select(system, class, entry_year, entry_age, age, yos, count, entry_salary)

saveRDS(salary_headcount, fs::path(drds, "salary_headcount.rds"))

```

## Create entrant profile

```{r}
#| label: entrant-profile

entrant_profile <- salary_headcount |>
  filter(entry_year == max(entry_year)) |>
  mutate(entrant_dist = count / sum(count), .by=c(system, class)) |>
  select(system, class, entry_age, start_sal=entry_salary, entrant_dist)

saveRDS(entrant_profile, fs::path(drds, "entrant_profile.rds"))

```

## Show the salary-headcount table

```{r}
#| label: show-salary-headcount-table

shc <- readRDS(fs::path(drds, "salary_headcount.rds"))

shc |> 
  # head() |> 
  DT::datatable(
    options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
    filter="top") |> 
  formatRound(columns = c("count"), digits=2) |>
  formatRound(columns = c("entry_salary"), digits=0)


# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```

## Show the new entrant profile table

```{r}
#| label: show-entrant-profile

ep <- readRDS(fs::path(drds, "entrant_profile.rds"))

ep |> 
  # head() |> 
  DT::datatable(
    options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
    filter="top") |> 
  formatRound(columns = c("start_sal"), digits=0) |> 
  formatPercentage(columns = c("entrant_dist"), digits=1)


# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```


## Comparison to Reason

```{r}
#| label: comp
#| eval: false


```
