# Termination rates

Get termination rate tables for FRS, from an Excel workbook (Florida FRS inputs.xlsx) that Reason created.


## TODO

Look for additional cleanup code that Reason has in other files and consolidate it here.

Weird eco withdrawal rate at 4 yos. Presumably ok.

deal with yos values. do we need to convert to single year?

## Setup

```{r}
#| label: setup
#| output: false

source("../_common_frs.R")

```

## Check excel tabs

```{r}
#| label: check

tabs <- excel_sheets(fullpath)
tabs |> str_subset("Withdrawal")
# read_excel(fullpath, sheet = "Withdrawal Rates")

```


## Get termination rates

```{r}
#| label: termination


wrtabs <- tabs |>
  str_subset("Withdrawal") |>
  setdiff("Withdrawal Rates")
wrtabs

f <- function(tab){
  print(tab)

  gender <- case_when(str_sub(tab, -4, -1) == "Male" ~ "male",
                      str_sub(tab, -6, -1) == "Female" ~ "female",
                      .default="all")

  class <- case_when(str_detect(tab, "Regular") ~ "regular",
                     str_detect(tab, "Special") ~ "special",
                     str_detect(tab, "Eco") ~ "eco",
                     str_detect(tab, "Eso") ~ "eso",
                     str_detect(tab, "Judges") ~ "judges",
                     str_detect(tab, "Sen Man") ~ "senior_management",
                     str_detect(tab, "Admin") ~ "admin",
                     .default = "ERROR")

  df1 <- read_excel(fullpath,
             sheet = tab,
             col_types = "text")

  df2 <- df1 |>
    mutate(system = "frs",
           class = class,
           gender = gender,
           yos = as.integer(yos)) |>
    pivot_longer(-c(system, class, gender, yos),
                 names_to = "age_group",
                 values_to = "term_rate")

  df2
}

f(wrtabs[1])

age_group_order <- c("under_25", "25_to_29", "30_to_34",
                     "35_to_44", "45_to_54", "over_55")

trates <- purrr::map(wrtabs, f) |>
  list_rbind() |>
  # percentage values read as character were not converted to decimal values
  mutate(term_rate=as.numeric(term_rate) / 100,
         # make age_group a factor so that it will sort in desired oder
         age_group = factor(age_group, levels = age_group_order)) |>
  select(system, class, gender, yos, age_group, term_rate) |>
  arrange(system, class, gender, yos, age_group)

skim(trates)
count(trates, class)
count(trates, gender)
count(trates, yos)
count(trates, age_group)

saveRDS(trates, path(drds, "termination_rates.rds"))

```


## Show the termination rates


```{r}
#| label: show-final-termination-rates

trates <- readRDS(fs::path(drds, "termination_rates.rds"))

trates |> 
  # head() |> 
  DT::datatable(
    options = list(scrollX = TRUE, scrollY = TRUE, paging = TRUE, pageLength = 20),
    filter="top") |> 
  formatPercentage(columns = c("term_rate"), digits=3)


# datatable(ptargets, options = list(scrollX = TRUE, scrollY = '400px', paging = TRUE))
# formatCurrency(columns = c("ptarget"), currency="", digits=0)
# formatPercentage(columns = c("salgrowth"), digits=2)
# formatRound(columns = c("cumprod_increase"), digits=4)

```

## Comparison to Reason

```{r}
#| label: comp
#| eval: false



# Truong's code, commented-out ----
# Termination rate tables

# regular_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Regular Male")
# regular_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Regular Female")
#
# special_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Special Male")
# special_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Special Female")
#
# admin_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Admin Male")
# admin_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Admin Female")
#
# eco_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Eco")
# eco_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Eco")
#
# eso_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Eso")
# eso_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Eso")
#
# judges_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Judges")
# judges_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Judges")
#
# senior_management_term_rate_male_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Sen Man Male")
# senior_management_term_rate_female_table_ <- read_excel(FileName, sheet = "Withdrawal Rate Sen Man Female")



```


